---
title: "Data Manipulation with dplyr"
author: "Yiying Wu"
date: "2023-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## tidyverse
```{r}
library(tidyverse)
```

## import data
Import FAS litters and pups.

```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```


## `select`
`select` is used to select columns
```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

`starts_with` select the cols starts with what you want:

```{r}
select(litters_data, group, starts_with("pups"))
```

specify the specify a range of columns to keep:
```{r}
select(litters_data, group:gd_of_birth)
```

You can also specify columns you’d like to remove:
```{r}
select(litters_data, -pups_survive)
```

You can rename variables as part of this process:
```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

`everything()`keep everything
`select` can select the cols in the order you write them
```{r}
select(litters_data, group, gd0_weight, everything())
```
`rename` function: change col names
```{r}
rename(litters_data,litter_id=litter_number)
```
`relocate` put the cols into the order you want
```{r}
relocate(litters_data,litter_id=litter_number)
```

`select` vs `pull`
```{r}
select(litters_data,group)
pull(litters_data,group)
```

## `filter`
get rid of rows using `filter()`

comparison operators (`>`, `>=`, `<`, `<=`, `==`, and `!=`).
```{r}
filter(litters_data, group =="Con7")
filter(litters_data, group =="Mod8")
filter(litters_data, group !="Con7")
filter(litters_data, gd0_weight>20)
filter(litters_data, gd0_weight<=20)
filter(litters_data, group=="Con7"|group=="Con8")
filter(litters_data, group %in% c("Con7","Con8"))
filter(litters_data, !(group=="Con7"))
filter(litters_data, group %in% c("Con7","Con8"),gd0_weight>20)
```

what about missing rows?
drop NAs in the whole dataset
```{r}
drop_na(litters_data)
```
drop NAs in specific col
```{r}
drop_na(litters_data,gd0_weight)
```



## `mutate`
this is used to add or change variables

`group=str_to_lower(variable1)` is going to convert the data in `variable1` into lowercases

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```


## `arrange`
arrange in ascending order by gd0_weight
```{r}
arrange(litters_data,gd0_weight)
```
arrange in alphabetical order by group then by gd0_weight
```{r}
arrange(litters_data,group,gd0_weight)
```

## pipes
Note that by default RStudio will insert the “traditional” pipe `%>%`, and you can update to the native in pipe through Global Preferences > Code > Use Native Pipe Operator. 
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") |> 
  janitor::clean_names() |> 
  select(-pups_survive) |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) |> 
  drop_na(wt_gain)
```

