---
title: "Data Manipulation with dplyr"
author: "Yiying Wu"
date: "2023-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## tidyverse
```{r}
library(tidyverse)
```

## import data
Import FAS litters and pups.

```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```


## `select`
`select` is used to select columns
```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

`starts_with` select the cols starts with what you want:

```{r}
select(litters_data, group, starts_with("pups"))
```

specify the specify a range of columns to keep:
```{r}
select(litters_data, group:gd_of_birth)
```

You can also specify columns you’d like to remove:
```{r}
select(litters_data, -pups_survive)
```

You can rename variables as part of this process:
```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

`everything()`keep everything
`select` can select the cols in the order you write them
```{r}
select(litters_data, group, gd0_weight, everything())
```
`rename` function: change col names
```{r}
rename(litters_data,litter_id=litter_number)
```
`relocate` put the cols into the order you want
```{r}
relocate(litters_data,litter_id=litter_number)
```



`select` vs `pull`
```{r}
select(litters_data,group)
pull(litters_data,group)
```

